
```{r}
library(tidyverse)
library(ggplot2)
library(lme4)
library(narray)
library(GADMTools)
library(outliers)
library(reshape)
library(vcd)
```

### 1.0
```{r 1.0}
shva <- read.table("https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/duryagin_ReductionRussian.txt", header = TRUE) 
```

### 1.1
```{r 1.1}
shva %>%
  ggplot(aes(f2, f1, fill = vowel)) +
  scale_y_continuous(trans = "reverse") + 
  scale_x_continuous(trans = "reverse") +
  geom_point(aes(col = vowel)) +
  labs(title = "f2 and f1 of reduced and stressed vowels",
       caption = "Data from Duryagin 2018") +
  guides(size = F, col = F, fill = F)
```
### 1.2
```{r 1.2}
shva %>%
  ggplot(aes(vowel, f1), label = vowel) +
  geom_boxplot(aes(vowel, f1, fill = vowel)) +
  coord_flip() +
  labs(title = "f1 distribution in each vowel",
       caption = "Data from Duryagin 2018",
       x = NULL) +
  guides(size = F, col = F, fill = F)
```
### 1.2
```{r 1.2}
shva %>%
  ggplot(aes(vowel, f2), label = vowel) +
  geom_boxplot(aes(vowel, f2, fill = vowel)) +
  coord_flip() +
  labs(title = "f2 distribution in each vowel",
       caption = "Data from Duryagin 2018",
       x = NULL) +
  guides(size = F, col = F, fill = F)
```
### 1.3
```{r 1.3}
shva %>%
  filter(vowel=="а") %>%
  mutate(shva, outlier = f1 > 4.5*IQR(f1))

```
### 1.4
```{r 1.4}
cor(shva$f1, shva$f2)
```
### The Pearson’s correlation of f1 and f2 for all data equals ~ -0.58.

### 1.5
```{r 1.5}
a_shva <- shva %>%
  filter(shva$vowel=="а")

cor(a_shva[,4:5], method = "pearson")
```

### 1.6
```{r 1.6}
f1_by_f2 <- lm(f1~f2, shva)
summary(f1_by_f2)
```

### 1.6.1
The regression formula is the following:
f1 = 1678,94 - 0,78f2

### 1.6.2
The adjusted R-squared equals 0.3319.

### 1.6.3
```{r 1.6.3}
shva$line1 <- predict(f1_by_f2)
shva$line1
```

```{r 1.6.3}
shva %>%
  ggplot(aes(f2, f1, fill = vowel)) +
  scale_y_continuous(trans = "reverse") + 
  scale_x_continuous(trans = "reverse") +
  geom_point(aes(col = vowel)) +
  labs(title = "f2 and f1 of the reduced and stressed vowels",
       caption = "Data from Duryagin 2018") +
  guides(size = F, col = F, fill = F) +
  geom_line(aes(f2, line1), color = "grey")
  
```
### 1.7
```{r 1.7}
f1_by_f2_v <- lmer(shva$f1 ~ shva$f2 + (1|shva$vowel))
summary(f1_by_f2_v)
```

### 1.7.1
The fixed effects formula is the following:
f1 = 489,32 + 0,06f1 

### 1.7.2
The variance for the intercept argument for vowel random effects equals 16741.

### 1.7.3
```{r 1.7.3}
shva$line2 <- predict(f1_by_f2_v)
shva$line2
```

```{r 1.7.3}
shva %>%
  ggplot(aes(f2, f1, fill = vowel)) +
  scale_y_continuous(trans = "reverse") + 
  scale_x_continuous(trans = "reverse") +
  geom_point(aes(col = vowel)) +
  labs(title = "f2 and f1 of the reduced and stressed vowels",
       caption = "Data from Duryagin 2018") +
  guides(size = F, col = F, fill = F) +
  geom_line(aes(f2, line2, color= vowel))
```

### 2.0
```{r 2.0}
elp <- read.csv("https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/ELP.csv", header = TRUE)
```

### 2.1
```{r 2.1}
l_s <- cor(elp$Length, elp$SUBTLWF)
s_m <- cor(elp$SUBTLWF, elp$Mean_RT)
l_m <- cor(elp$Length, elp$Mean_RT)
l_s
s_m
l_m
```
"Length" and "Mean_RT" have the highest Pearson`s correlation ~0.53.

### 2.2
```{r 2.2}
elp %>%
  group_by(POS) %>%
  ggplot(aes(log(SUBTLWF), Mean_RT), color = Length, xlog = FALSE) +
  geom_point(aes(color = Length)) +
  scale_color_continuous(low = "lightblue", high = "red") +
  facet_wrap(~POS) +
  theme_bw() +
  #geom_point(aes(Length), col = "red", position = "dodge") +
  labs(caption = "data from (Balota et al. 2007)",
       x = "SUBTLWF",
       y = "Mean_RT")
```

### 2.3
```{r 2.3}
mult_regr <- lm(Mean_RT~log(SUBTLWF)+POS, data = elp)
summary(mult_regr)
```

### 2.3.1
Mean_RT = 786.18 - 37.57*log(SUBTLWF)-12.53*NN-42.8*VB

### 2.3.2
The Adjusted R-squared equals 0.3385

### 2.3.4
```{r 2.3.4}
elp$line3 <- predict(lm(elp$Mean_RT~log(elp$SUBTLWF)))
log_s <- log(elp$SUBTLWF)
```

```{r 2.3.4}
elp %>%
  group_by(POS) %>%
  ggplot(aes(log_s, Mean_RT), color = Length, xlog = FALSE) +
  geom_point(aes(color = Length)) +
  scale_color_continuous(low = "lightblue", high = "red") +
  theme_bw() +
  geom_line(aes(y = line3)) +
  labs(caption = "data from (Balota et al. 2007)",
       x = "SUBTLWF",
       y = "Mean_RT")
```

### 2.4
```{r 2.4}

mult_rand <- lmer(Mean_RT~log_s+(1|POS), data = elp)
summary(mult_rand)
```

### 2.4.1
The fixed effects formula:
Mean_RT = 767.71 - 37.67*log(SUBTLWF)

### 2.4.2
The variance for intercept argument for POS random effects equals 414.4

### 2.4.3
```{r 2.4.3}
elp$line4 <- predict(mult_rand)
```

```{r 2.4.3}
elp %>%
  group_by(POS) %>%
  ggplot(aes(log(SUBTLWF), Mean_RT), color = Length) +
  geom_point(aes(color = POS)) +
  facet_wrap(~POS) +
  geom_line(aes(y = line4)) +
  theme_bw() +
  labs(caption = "data from (Balota et al. 2007)",
       x = "SUBTLWF",
       y = "Mean_RT")
```

### 3.0
```{r 3.0}
d_caus <- read.csv("https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/dutch_causatives.csv", header = TRUE)
```

### 3.1
```{r 3.1}
library(car)
d_caus%>%
  count(Aux)

```
```{r 3.1}
d_caus%>%
  count(CrSem)
```
```{r 3.1}
d_caus%>%
  count(CdEvSem)
```
```{r 3.1}
d_caus%>%
  count(CeSynt)
```
```{r 3.1}
d_caus%>%
  count(EPTrans)
```
```{r 3.1}
d_caus%>%
  count(Country)
```
```{r 3.1}
d_caus%>%
  count(Domain)
```
  

```{r 3.1.1}
aux_crsem <- cbind(c(85, 415), c(408, 92))
aux_crsem

chisq.test(aux_crsem)$expected
chisq.test(aux_crsem)

```
 1) The expected values are higher than 5, so, we can perform the chi-squared test.
 2) As the p-value is under 0.05, we can state being 95% sure that the assosiation between Aux and CrSem (the semantic class of the Causer) is statistically significant.

```{r 3.1.2}
aux_cdevsem <- cbind(c(85, 415), c(101, 399))
aux_cdevsem

chisq.test(aux_cdevsem)$expected
chisq.test(aux_cdevsem)
```
 The p-value is higher than 0.05, so, the assosiation between Aux and CrEvSem (the semantic class of the Causee) is NOT statistically significant.

 In this case we are recommended to check the Fisher exact test:
```{r 3.1.2}
fisher.test(aux_cdevsem)
```
 The p-value is also higher than 0.05. For this reason the assosiation between Aux and CrEvSem (the semantic class of the Causee) is NOT statistically significant.


```{r 3.1.4}
aux_country <- cbind(c(85, 415), c(220, 280))
aux_country

chisq.test(aux_country)$expected
chisq.test(aux_country)
```
 As the p-value is under 0.05, we can state being 95% sure that the assosiation between Aux and Country variables is statistically significant.

### 3.2
```{r 3.2}
aux_trans <-cbind(c(85, 415), c(239, 261))
aux_trans

chisq.test(aux_trans)
```
 As the p-value is under 0.05, we can state being 95% sure that the the variables Aux and EPTrans are not independent.

### 3.3
```{r 3.3}
chisq.test(aux_trans)$expected
```

### 3.4
```{r 3.4}
fisher.test(aux_trans)
```

 The odds ratio equals 0.22 which means that EPTrans odds are 5 times higher than Aux odds.

### 3.5
```{r 3.5}
assocstats(aux_trans)
```

 The Phi-Coefficient, Cramer`s V equal 0.329 which tells us about a moderate effect size.

### 3.6
```{r 3.6}
```
 We have found a significant association between variables Aux and CrSem, Aux and CrEvSem, Aux and EPTrans, Aux and Country (p < 0.001).
 The odds of aux_trans variable were 5 times higher in EPTrans than in Aux. Effect size is  medium (Cramer's V = 0.329).

### 3.7
```{r 3.7}
mosaic(~Aux+EPTrans, data = d_caus, shade = TRUE)
```

### 3.9
 Basing on the data of Dutch causative constructions we have checked the interconnection of several grammatical and semantic elements. We assumed that there is the correlation between:
1)	auxiliary type (Aux)  and semantic class of the causer (CrSem),
2)	auxiliary type (Aux)  and semantic class of the causee (CdEvSem),
3)	auxiliary type (Aux) and the transitivity/intransitivity (EPTrans),
4)	auxiliary type and the country of origin.
 To analyze the hypotheses we have performed the chi squared test of independence (to be able to count it, we counted the observed frequencies for each variable and got the contingency tables).
 As a result, we found out the dependency between the variables in the 3 cases. The high p-value of the model does not let us state the correlation between auxiliary type and semantic class of the cause. 
 Our main attention was stuck to the auxiliary type and the transitivity/intransitivity relation. The low p-value (<0.05) lets us state that the variables are somehow interdependent. Having counted the odds ratio, we found out that the EPTrans are five times higher than the Aux odds. The effect size is  medium (Cramer's V = 0.329).
